### 2022-04-23 outputs/paper/paper.Rmd
Adding starter folder to repository

---
title: "Salaries and Rising Police Force Budgets"
subtitle: "An Exploratory Analysis on Expenditures of Toronto Police Force"
author: 
  - Meha Grewal
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "In this paper we will explore the budget of the Toronto Police Force and examine the implications of their fiscal responsibilities. Through the understanding of the central influences of the police force, we can determine how to effectively distribute resources and funding. These findings provide rationale for policy decisions."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

```{r}
data <- read.csv("/Users/meha/Downloads/TPS Budget 2021.csv")
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

The Toronto Police Force (TPF) are supposed to be the upholders of law and order. They provide a sense of safety to the community and work to prevent and solve criminal cases. The budget for the the TPF is highly significant as it allows them the ability to safely govern the city. To identify the most influential aspects of the TPF budget and thus the parts of policing that they deem as priority, we will analyze the relevant variables and model them. We explore their expenditures and fiscal responsibilities for 2021 and its implications in providing safety for residents of Toronto.

Conducting an exploratory analysis on the budget for TPF would indicate areas that the government prioritizes. Utilizing the dollar amount as the response variable, we can conclude the largest predictors for police force expenditures and its implications. This information can be applied by government officials and policy makers to minimize government spending and allocate finances to areas that most necessitate it.

Beginning with a general overview of the variables provided in the data, we visually identified where majority of the budget is spent on. Through a multiple linear regression model, we were able to determine the most statistically significant variables. This indicates what portions of the police force have the most influence over the budget and the priorities of the force. By analyzing each predictor to understand its importance, we then considered the broader societal implications. The linear regression indicates how finances are allocated and which districts and units relieve the largest portions of the budget. By determining what districts and categories within the police force require the most funding, we can evaluate the future of the force. This analysis provides the determining factors of the fiscal responsibilities for the police and showcases what areas require more resources. Policy decisions influence the acceptable expenditures of the TPF and this paper aims to understand the ramifications of the budget distribution.

The data set provided by the city of Toronto showcases information regarding actual expenditures from the police force in 2021. By examining the relevant summary statistics and plots, we are able to identify key predictors for our model. We then model the data using the amount for the budget as the response variable and conduct a multiple linear regression. This was done to predict the budget and the crucial components. Our discussion includes the understanding of our model and how it was assembled. Lastly, we consider the social implications of the model and how the budget has an impact on safety in Toronto.

# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional details

\newpage

# References